---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
# INTRODUCTION
It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. 
These type of devices are part of the "quantified self" movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. 
But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.

This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

The variables included in this dataset are:

- steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)
- date: The date on which the measurement was taken in YYYY-MM-DD format
- interval: Identifier for the 5-minute interval in which measurement was taken

The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.

## 1 - Loading and preprocessing the data

* Loading the Data

```{r}
getwd()
library(knitr)
library(markdown)
library(rmarkdown)
library("Hmisc")
library(dplyr)

```

```{r}
dataset <- tempfile()
download.file("http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", dataset)
dataset <- unz(dataset, "activity.csv")
activity <- read.csv(file = "activity.csv", header = TRUE, sep = ",")
unlink(dataset)
rm(dataset)
```

```{r, echo=TRUE}
names(activity)
```

```{r, echo= TRUE}
str(activity)
```


* Transforming the data for suitable analysis

```{r}
## activity$date <- as.Date(activity$date)
## str(activity)
## activity <- activity[which(activity$steps!= "NA"), ]
## str(activity)
```



## 2 - What is mean total number of steps taken per day?

* Calculate the sum of steps taken by day

```{r, echo= TRUE}
sumSteps <- aggregate(steps ~ date, data = activity,  FUN="sum")
sumSteps
```

* Histogram of the total number of steps taken each day.

```{r, echo = TRUE}
hist(sumSteps$steps, col = 'blue',  xlab = 'Total Steps per Day', main = "Histogram of Activities")
dev.off()
```

* Calculate the mean and median of the total number of steps taken per day 

* Total Mean Steps taken per day
```{r, echo = TRUE}
meanStepsPerDay <- aggregate(steps ~ date, data = activity, FUN = "mean")
meanStepsPerDay
totalMeanSteps <- mean(sumSteps$steps)
totalMeanSteps
```

* Total Median Steps 
```{r, echo = TRUE}
totalMedianSteps <- median(sumSteps$steps)
totalMedianSteps
```



## 3 - What is the average daily activity pattern?

* Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r, echo=TRUE}
averageSteps <- aggregate(steps~interval, data=activity, FUN="mean")
averageSteps
plot(x = averageSteps$interval, y = averageSteps$steps, type = "l", lwd=2, xlab = "Intervals",
     ylab = "Average number of steps", main = "Average Daily activity Pattern")
dev.off()
```


* Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of step

```{r, echo=TRUE}
averageSteps[which.max(averageSteps$steps),]
averageSteps[which.max(averageSteps$steps),]$interval

```


## 4 - Imputing missing values - STRATEGY

* 4.1 = Total number of missing values (total rows with NAs) 

```{r, echo=TUE}
sum(is.na(activity))
```

* 4.2 = Filling the the missing data with average using a new Dataset

```{r, echo=TRUE}


```


* 4.3 = Histogram of the total number of steps taken each day after missing values are imputed.

```{r, echo=TRUE}
hist(filledsteps,breaks=61,main="Steps Per Day By Frequency",xlab="Daily Steps")
```

* 4.4- Mean and Median

```{r, echo=TRUE}

```

############################ 
```{r}

```







## 5 - Are there differences in activity patterns between weekdays and weekends?

```{r}

```
